Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2024-02-19 paper/main.tex
camera ready submitted

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[FSE '24]{Foundations of Software Engineering}{July 2024}{Brazil}

%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

\author{Sourav Deb}
\affiliation{\institution{Northern Arizona University}
\country{United States}}

\author{Kush Jain}
\affiliation{\institution{Carnegie Mellon University}
\country{United States}}

\author{Rijnard van Tonder}
\affiliation{\institution{Mysten Labs}
\country{United States}}

\author{Claire Le Goues}
\affiliation{\institution{Carnegie Mellon University}
\country{United States}}

\author{Alex Groce}
\affiliation{\institution{Northern Arizona University}
\country{United States}}

@@

%\keywords{Software Testing, Mutants, Mutation Score, Equivalence}

@@

Mutation testing, though introduced in the late
1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
has long been a topic of academic interest.  With the recent adoption of mutation
testing by bellwether software industry companies including
Google~\cite{GoogleMut}, Meta~\cite{BellerFacebookMutation}, and
Amazon~\cite{AmazonMut}, interest in using mutation testing for
real-world projects has grown widely in practice. 
%
Indeed, interest in mutation testing has become widespread enough in the open
source community that a popular (more than 300 stars on GitHub)
repository, ``Awesome Mutation Testing,'' lists more than 40 tools,
covering almost twenty languages.\footnote{\url{https://github.com/theofidry/awesome-mutation-testing}}  Some of
these tools are the products of academic research; others are
essentially the hobby projects of developers interested in enabling
mutation testing for their favorite language.

@@

We found only one tool that allowed single-file mutation and worked
reliably for C++.  Dextool generally produced fewer mutants than
either mode of our tool, and for some projects produced an extremely
small number of mutants.  For example, for the {\tt
  tiny-differentiable-simulator} project, Dextool only produced a
total of 61 mutants, and gave a much higher mutation score than our
two modes.  Our tools produced more than 600 additional unkilled
mutants, and none of the inspected mutants were equivalent.
Similarly, Dextool only found 2 unkilled mutants for the {\tt
  libgraphqlparser} project.  We doubt that the set of test weaknesses
in the code is adequately represented by such a small number of
mutants when our tools were able to produce significantly more
unkilled mutants.  Even assuming Dextool's mutants are never
equivalent (possible though not necessarily likely) and supposing that a third of our
generated mutants are equivalent, it seems likely that Dextool is
missing important mutants in some cases.  Overall, mean mutation
scores were similar for Dextool and our tool, while
median mutation scores were much lower for our tool, which produced
many more unkilled mutants for some projects.  While our equivalence
rate was somewhat high for one project, it was generally good
compared to results in the literature.  For C++, there seems to
be no reason to suspect that the universal approach is producing
sub-par mutation testing.

@@

>>>>

%%% The following is specific to FSE '24 and the paper
%%% 'Syntax Is All You Need: A Universal-Language Approach to Mutant Generation'
%%% by Sourav Deb, Kush Jain, Rijnard van Tonder, Claire Le Goues, and Alex Groce.
%%%
\setcopyright{acmlicensed}
\acmDOI{10.1145/3643756}
\acmYear{2024}
\copyrightyear{2024}
\acmSubmissionID{fse24main-p529-p}
\acmJournal{PACMSE}
\acmVolume{1}
\acmNumber{FSE}
\acmArticle{30}
\acmMonth{7}
\received{2023-09-27}
\received[accepted]{2024-01-23}

@@

\author{Sourav Deb}
\affiliation{\institution{Northern Arizona University}
  \city{Flagstaff, AZ}
\country{United States}}

\author{Kush Jain}
\affiliation{\institution{Carnegie Mellon University}
  \city{Pittsburgh, PA}
\country{United States}}

\author{Rijnard van Tonder}
\affiliation{\institution{Mysten Labs}
  \city{Palo Alto, CA}
\country{United States}}

\author{Claire Le Goues}
\affiliation{\institution{Carnegie Mellon University}
  \city{Pittsburgh, PA}  
\country{United States}}

\author{Alex Groce}
\affiliation{\institution{Northern Arizona University}
  \city{Flagstaff, AZ}
\country{United States}}

@@

\keywords{Software Testing, Mutants, Mutation Generation}

@@

Mutation testing, though introduced in the late
1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
has long been a topic of academic interest.  With the recent adoption of mutation
testing by bellwether software industry companies including
Google~\cite{GoogleMut}, Meta~\cite{BellerFacebookMutation}, and
Amazon~\cite{AmazonMut}, interest in using mutation testing for
real-world projects has grown widely in practice.

Indeed, interest in mutation testing has become widespread enough in the open
source community that a popular (more than 300 stars on GitHub)
repository, ``Awesome Mutation Testing,'' lists more than 40 tools,
covering almost twenty languages.\footnote{\url{https://github.com/theofidry/awesome-mutation-testing}}  Some of
these tools are the products of academic research; others are
essentially the hobby projects of developers interested in enabling
mutation testing for their favorite language.

@@

We found only one tool that allowed single-file mutation and worked
reliably for C++.  Dextool generally produced fewer mutants than
either mode of our tool, and for some projects produced an extremely
small number of mutants.  E.g., for the {\tt
  tiny-differentiable-simulator} project, Dextool only produced a
total of 61 mutants, and gave a much higher mutation score than our
two modes.  Our tools produced more than 600 additional unkilled
mutants, and none of the inspected mutants were equivalent.
Similarly, Dextool only found 2 unkilled mutants for the {\tt
  libgraphqlparser} project.  We doubt that the set of test weaknesses
in the code is adequately represented by such a small number of
mutants when our tools were able to produce significantly more
unkilled mutants.  Even assuming Dextool's mutants are never
equivalent (possible though not necessarily likely) and supposing that a third of our
generated mutants are equivalent, it seems likely that Dextool is
missing important mutants in some cases.  Overall, mean mutation
scores were similar for Dextool and our tool, while
median mutation scores were much lower for our tool, which produced
many more unkilled mutants for some projects.  While our equivalence
rate was somewhat high for one project, it was generally good
compared to results in the literature.  For C++, there seems to
be no reason to suspect that the universal approach is producing
sub-par mutation testing.

@@

\section{Acknowledgments}

A portion of this work was funded by the National Science Foundation 
under NSF awards CCF-2129388, CCF-1910067, and CCF-2129446.  We also 
thank all contributors to the Universalmutator and Comby projects.
